High angular resolution dust density and temperature maps from \Herschel\ imaging data

### 2012-08-05 documents/herschel_dust.tex
tiny typo
cross validation explained

\title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
\author{
  David~W.~Hogg\altaffilmark{1,2,3},
  Dustin~Lang\altaffilmark{4,5},
  Brent~Groves\altaffilmark{3},
  Hans-Walter~Rix\altaffilmark{3},
  Karin~Sandstrom\altaffilmark{3}}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
\altaffiltext{2}{Center for Cosmology and Particle Physics, New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
\altaffiltext{4}{Princeton University Observatory}
\altaffiltext{5}{Carnegie Mellon University}

@@

In non-parametric models like this---even when we only want point
estimates or optima of an objective function---we cannot avoid using
prior information to regularize the fit.  The priors we choose are
``smoothness'' priors inspired by Gaussian processes.  They penalize
the sum of squares of differences between neighboring pixels.
\begin{eqnarray}\displaystyle
-2\,\ln p(\Omega_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_\Omega\,\left[\Omega_k - \Omega_{k'}\right]^2
\\
-2\,\ln p(T_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_T\,\left[T_k - T_{k'}\right]^2
\\
-2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta - 2\right]^2
                    + \sum_{\mathrm{neighbors}~k'} \epsilon_\beta\,\left[\beta_k - \beta_{k'}\right]^2
\\
\alpha &\equiv& \left\{\epsilon_\Omega, \epsilon_T, \epsilon_2, \epsilon_\beta\right\}
\quad ,
\end{eqnarray}
where again the ``$\approx$'' symbols are used as a reminder that
these priors are not properly normalized as written, the sum over
``neighbors'' $k'$ is a sum, for each pixel, over its four nearest
neighbors, the prior ``strength'' or information is set by the list
$\alpha$ of hyper-parameters $\{\epsilon_\Omega, \epsilon_T,
\epsilon_2, \epsilon_\beta\}$, and the emission parameter grid
$\beta_k$ is given the additional prior that we expect values around
2.

...The hyperparameters are set by cross-validation...

>>>>

\title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
\author{
  Dustin~Lang\altaffilmark{4,5},
  David~W.~Hogg\altaffilmark{1,2,3},
  Brent~Groves\altaffilmark{3},
  Hans-Walter~Rix\altaffilmark{3},
  Karin~Sandstrom\altaffilmark{3}}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
\altaffiltext{2}{Center for Cosmology and Particle Physics, New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
\altaffiltext{4}{Princeton University Observatory}
\altaffiltext{5}{Carnegie Mellon University}

@@

In non-parametric models like this---even when we only want point
estimates or optima of an objective function---we cannot avoid using
prior information to regularize the fit.  There are no extremely good
priors to put on images, and certainly not images of interstellar
dust.  For this reason we choose simply ``smoothness'' priors inspired
by Gaussian processes.  They penalize the sum of squares of
differences between neighboring pixels.
\begin{eqnarray}\displaystyle
-2\,\ln p(\Omega_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_\Omega\,\left[\Omega_k - \Omega_{k'}\right]^2
\\
-2\,\ln p(T_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_T\,\left[T_k - T_{k'}\right]^2
\\
-2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta_k - 2\right]^2
                    + \sum_{\mathrm{neighbors}~k'} \epsilon_\beta\,\left[\beta_k - \beta_{k'}\right]^2
\\
\alpha &\equiv& \left\{\epsilon_\Omega, \epsilon_T, \epsilon_2, \epsilon_\beta\right\}
\quad ,
\end{eqnarray}
where again the ``$\approx$'' symbols are used as a reminder that
these priors are not properly normalized as written, the sum over
``neighbors'' $k'$ is a sum, for each pixel, over its four nearest
neighbors, the prior ``strength'' or information is set by the list
$\alpha$ of hyper-parameters $\{\epsilon_\Omega, \epsilon_T,
\epsilon_2, \epsilon_\beta\}$, and the emission parameter grid
$\beta_k$ is given the additional prior that we expect values around
2.

The posterior probability is proportional to the likelihood times the
prior; its logarithm is sum sum of the log likelihood and the log
prior.  Dust map inference---the determination of the solid-angle
parameters $\Omega_k$, the temperature parameters $T_k$, and the
emissivity parameters $\beta_k$ proceeds by optimization of the log
posterior probability.  The parameters so derived---maximum \foreign{a
  posteriori} or MAP---are taken to be the ``best-fit'' parameters in
what follows.  The MAP dust-map parameters depend on the data, the
spatial grid $\theta_k$, the data $d_n$, and the data noise variances
$\sigma_n$.

The hyperparameters $\alpha$---parameters of the priors---set the
strictness or information content of the priors.  We want to give
these hyperparameters values that have been dervied objectively from
the data, so that the collection of \emph{all} data informs the prior
probability for the parameters $\Omega_k, T_k, \beta_k$ affecting any
small part of the data.  The best methods for such hyperparameter
inference is hierarchical Bayesian inference, which is too
computationally costly for the present effort.  The approximation to
hierarchical inference adopted here is that of leave-some-out
cross-validation: We drop randomly chosen small subsets of the data,
perform the inference of the dust-map parameters $\Omega_k, T_k,
\beta_k$ using the remaining (the non-left-out) data, predict the
left-out data, and base the hyperparameter settings on the quality of
the predictions.

In detail, this works by choosing a random integer $1\leq s\leq S$
(where $S$ is the number of leave-some-out trials, $S=8$ in what
follows) for every data pixel $n$.  For each value of $s$---say
3---MAP values for the dust-map parameters are obtained given all the
data \emph{except} that marked with integer $s=3$.  The data marked
with integer $s=3$ are predicted by this model, and a prediction
$\chi^2$ value is computed for the left-out $s=3$ data.  This is
repeated for all values of $s$ and all 8 prediction $\chi^2$ values
are summed.  This total prediction leave-some-out $\chi^2$ is a
function of the hyperparameters; the hyperparameters $\alpha$ at which
it is minimized are chosen as the best hyperparameters.

One extremely valuable side-effect of the leave-one-out trials is that
they also provide uncertainty estimates on the dust-map parameters.
The leave-some-out trials are effectively also a set of jackknife
trials.  The variance (mean squared deviation) in the map parameters
across the $S$ leave-some-out trials is multiplied by (HOGG: FORMULA
HERE) to give a conservative empirical (bootstrap-like) estimate of
the variance (mean squared error) in the map parameters.
